### 2022-11-18 article_1/solution_par_table.tex

\begin{table}[htbp]
\caption{Positioning and magnetization parameters recovered with the inversion.}
\label{tab:InversionTab}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{ccccccc}
\hline   
\rowcolor[HTML]{FFFFFF} 
\cellcolor[HTML]{FFFFFF}                         & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Center coordinates}                    & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Magnetization}     \\
\rowcolor[HTML]{FFFFFF} 
\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}Sphere} & Xc ($\mu m$) & Yc ($\mu m$) & Zc ($\mu m$) & m ($A \cdot m^2$) & D (\textdegree) & I (\textdegree) \\  \hline   
1                                                & 250.03           & 249.99           & 5.33           & 8.75e-15  $\pm$  2.30e-17           & -140.04  $\pm$  0.19           & -29.76  $\pm$  0.13           \\
\rowcolor[HTML]{FFFFFF} 
2                                                & 499.96           & 500.05           & 7.70           & 7.60e-15  $\pm$  2.90e-17           & 1.17  $\pm$  0.61           & 62.50  $\pm$  0.27           \\
3                                                & 750.12           & 750.04           & 8.47           & 6.19e-15  $\pm$  3.52e-17           & -71.88  $\pm$  0.60           & -50.02  $\pm$  0.35           \\
\rowcolor[HTML]{FFFFFF} 
4                                                & 200.02           & 799.73           & 10.08           & 6.73e-15  $\pm$  5.05e-17           & 124.28  $\pm$  0.48           & 22.78  $\pm$  0.34           \\ \hline                 
\end{tabular}
}
\end{table}