### 2022-11-24 article_1/solution_par_table.tex

\begin{table}[htbp]
\caption{Positioning and magnetization parameters recovered with the inversion.}
\label{tab:InversionTab}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{ccccccc}
\hline   
\rowcolor[HTML]{FFFFFF} 
\cellcolor[HTML]{FFFFFF}                         & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Center coordinates}                    & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Magnetization}     \\
\rowcolor[HTML]{FFFFFF} 
\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}Sphere} & Xc ($\mu m$) & Yc ($\mu m$) & Zc ($\mu m$) & m ($A \cdot m^2$) & D (\textdegree) & I (\textdegree) \\  \hline   
1                                                & 250.07           & 250.04           & 5.36           & 8.83e-15  $\pm$  2.32e-17           & -140.03  $\pm$  0.18           & -29.45  $\pm$  0.13           \\
\rowcolor[HTML]{FFFFFF} 
2                                                & 500.00           & 500.01           & 7.71           & 7.57e-15  $\pm$  2.91e-17           & 0.23  $\pm$  0.60           & 61.79  $\pm$  0.27           \\
3                                                & 749.92           & 749.98           & 8.38           & 6.10e-15  $\pm$  3.46e-17           & -69.93  $\pm$  0.61           & -50.50  $\pm$  0.35           \\
\rowcolor[HTML]{FFFFFF} 
4                                                & 199.94           & 800.21           & 10.10           & 6.78e-15  $\pm$  5.09e-17           & 125.09  $\pm$  0.48           & 21.02  $\pm$  0.33           \\ \hline                 
\end{tabular}
}
\end{table}