### 2022-12-01 article_1/input_par_table.tex

\begin{table}[htbp]
\caption{Initial position and magnetization parameters for each spherical source.}
\label{tab:ModelTab}
\resizebox{1.0\textwidth}{!}{
\tiny
\begin{tabular}{ccccccc}
\hline   
\rowcolor[HTML]{FFFFFF} 
\cellcolor[HTML]{FFFFFF}                         & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Center coordinates}                    & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Magnetization}     \\
\rowcolor[HTML]{FFFFFF} 
\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}Sphere} & Xc ($\mu m$) & Yc ($\mu m$) & Zc ($\mu m$) & m ($A \cdot m^2$) & D (\textdegree) & I (\textdegree) \\  \hline   
1                                                & 250.00                       & 250.00                       & 5.30                       & 8.70e-15               & -140.00                        & -30.00                 \\
\rowcolor[HTML]{FFFFFF} 
2                                                & 500.00                       & 500.00                       & 7.75                       & 7.63e-15               & -70.00                        & -50.00                 \\
3                                                & 750.00                       & 750.00                       & 8.50                       & 6.21e-15               & 10.00                        & 62.00                 \\
\rowcolor[HTML]{FFFFFF} 
4                                                & 200.00                       & 800.00                       & 10.00                       & 4.97e-15               & 125.00                        & 22.00                  \\ \hline                 
\end{tabular}
}
\end{table}