### 2022-12-02 article/solution_par_table.tex

\begin{table}[htbp]
\caption{Positioning and magnetization parameters recovered with the inversion.}
\label{tab:InversionTab}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{ccccccc}
\hline   
\rowcolor[HTML]{FFFFFF} 
\cellcolor[HTML]{FFFFFF}                         & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Center coordinates}                    & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Magnetization}     \\
\rowcolor[HTML]{FFFFFF} 
\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}Sphere} & Xc ($\mu m$) & Yc ($\mu m$) & Zc ($\mu m$) & m ($A \cdot m^2$) & D (\textdegree) & I (\textdegree) \\  \hline   
1                                                & 250.03           & 249.94           & 5.37           & 8.80e-15  $\pm$  1.39e-17           & -140.22  $\pm$  0.11           & -30.04  $\pm$  0.08           \\
\rowcolor[HTML]{FFFFFF} 
2                                                & 499.96           & 500.00           & 7.76           & 7.65e-15  $\pm$  1.89e-17           & -69.42  $\pm$  0.26           & -49.85  $\pm$  0.15           \\
3                                                & 749.98           & 749.97           & 8.56           & 6.29e-15  $\pm$  1.98e-17           & 9.06  $\pm$  0.50           & 62.34  $\pm$  0.22           \\
\rowcolor[HTML]{FFFFFF} 
4                                                & 200.10           & 800.12           & 9.99           & 4.92e-15  $\pm$  3.00e-17           & 124.63  $\pm$  0.39           & 21.76  $\pm$  0.27           \\ \hline                 
\end{tabular}
}
\end{table}