### 2022-12-19 article_1/solution_par_table.tex
Delete unused images

\begin{table}[htbp]
\caption{Positioning and magnetization parameters recovered with the inversion.}
\label{tab:InversionTab}
\resizebox{1.0\textwidth}{!}{
\begin{tabular}{ccccccc}
\hline   
\rowcolor[HTML]{FFFFFF} 
\cellcolor[HTML]{FFFFFF}                         & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Center coordinates}                    & \multicolumn{3}{c}{\cellcolor[HTML]{FFFFFF}Magnetization}     \\
\rowcolor[HTML]{FFFFFF} 
\multirow{-2}{*}{\cellcolor[HTML]{FFFFFF}Sphere} & Xc ($\mu m$) & Yc ($\mu m$) & Zc ($\mu m$) & m ($A \cdot m^2$) & D (\textdegree) & I (\textdegree) \\  \hline   
1                                                & 250.28           & 250.19           & 5.38           & 8.82e-15  $\pm$  1.39e-17           & -140.22  $\pm$  0.11           & -30.04  $\pm$  0.08           \\
\rowcolor[HTML]{FFFFFF} 
2                                                & 500.46           & 500.50           & 7.77           & 7.67e-15  $\pm$  1.90e-17           & -69.42  $\pm$  0.26           & -49.85  $\pm$  0.15           \\
3                                                & 750.73           & 750.72           & 8.57           & 6.31e-15  $\pm$  1.99e-17           & 9.06  $\pm$  0.50           & 62.34  $\pm$  0.22           \\
\rowcolor[HTML]{FFFFFF} 
4                                                & 200.30           & 800.92           & 10.00           & 4.94e-15  $\pm$  3.01e-17           & 124.63  $\pm$  0.39           & 21.76  $\pm$  0.27           \\ \hline                 
\end{tabular}
}
\end{table}