### 2023-03-14 paper/content.tex
Add changes to the paper manuscript

\section{Introduction}

Several branches of Earth Sciences have demonstrated the importance of the
``spatiality" of the data on a microscopic scale, mainly in geochemistry and
geochronology applications where it is possible to perform punctual analysis
and compositional maps, allowing significant advances in the understanding of
igneous, metamorphic and sedimentary processes \citep{Verberne2020, Barnes2019,
Davidson2007}. Classical paleomagnetic techniques, on the other hand, consist
in analyzing bulk samples, where the magnetic signal of a single specimen is
the result of the sum of moments of a large assembly of ferromagnetic grains
\citep{Dunlop1997}. Typically, a standard paleomagnetic sample of
approximately \qty{10}{\cm\cubed} would contain hundreds of thousands
to millions of magnetic particles with sizes varying from magnetically stable
single-domain (SD) and vortex state grains (also called pseudo-single domain,
PSD) with sizes below \qty{1}{\um}, to large ($\gg \qty{1}{\um}$) grains with
multi-domain (MD) magnetic structures, which are less stable magnetic recorders
\citep{Berndt2016}. These large MD grains usually conceal the signal of the SD
and PSD grains, and techniques of step-wise thermal and magnetic treatments are
needed to unveil the more stable and reliable magnetic record
\citep{Tauxe2018}. Recently, magnetic microscopy techniques opened the
possibility of performing magnetic maps at the micro-scale and recovering the
magnetization of each grain, therefore enabling the separate analysis of stable
and unstable magnetic particles \citep{DeGroot2018, Lima2014, Weiss2007,
DeGroot2014}.

In order to apply magnetic microscopy to paleomagnetic studies, it is necessary
to recover from the magnetic images a large amount of individual magnetic
moments, corresponding to at least tens of thousands of stable fine-grained
grains ($< \qty{1}{\um}$), in order to provide statistical significance to the
remanence vector \citep[e.g., ][]{Berndt2016}. Nowadays, with the development of
magnetic microscopy techniques, this task is no longer limited by the
resolution of magnetic microscopes \citep{Fu2020, Weiss2007, DeGroot2018,
Glenn2017, Lima2014}, but essentially by the intrinsic problem presented by the
ambiguity in the inversion of potential field data \citep{Barbosa2011,
DeGroot2021, Oliveira2015Estimation}, and ultimately by the lack of a fast and
automated way to recover such a large number of individual magnetic moments
from a set of magnetic images \citep{CortesOrtuno2022, Lima2013, Lima2009}. A
solution to the non-uniqueness of magnetic moment inversion is to add
independent prior information, such as the position of the ferromagnetic
particles \citep{Fabian2019}. This can be obtained, for example, from X-ray
computed tomography \citep[microCT; ][]{Fabian2019, DeGroot2021, DeGroot2018}.
Nonetheless, the standard microCT techniques do not provide adequate resolution
to resolve the finer and more stable magnetic grains \citep{CortesOrtuno2022,
DeGroot2021}, whereas other more sophisticated techniques such as ptychographic
X-ray tomography \citep[e.g., ][]{Maldanis2020} are not readily available and
too time-consuming to be routinely used in paleomagnetic studies.

Another route to be explored in the inversion of magnetic microscopy images is
to obtain all the information, i.e. the magnetic moment and the position of the
sources, from the magnetic data itself \citep[e.g., ][]{Fu2020}. For that, we can
explore the techniques developed in exploration geophysics, in spite of the
differences between aeromagnetic surveys and magnetic microscopy
\citep{Lima2013}. Magnetic microscopy images commonly show the combined signal
of multiple magnetic particles and can vary greatly in wavelength, strength,
and spatial separation, depending on the NRM and location of each particle. We
usually assume that the signal measured by the magnetic microscope is the
vertical component of the magnetic induction vector ($b_z$), the measurements
are performed on a regular grid with evenly spaced grid points and at a
constant height, and the data are contaminated with random Gaussian noise and
long-wavelength noise (akin to a regional signal in aeromagnetic data). Here,
we provide a methodological routine to retrieve the individual magnetic moment
of ferromagnetic grains in magnetic microscopy images following the approach
devised by \citet{Oliveira2015Estimation} for the interpretation of
aeromagnetic anomalies. The method we propose allows one to quickly and
semi-automatically estimate the individual magnetic moment vector of the stable
magnetic carriers, making use of only the magnetic images themselves and an
assumption of approximately dipolar sources. If used on a large scale, the
method provides the means to scan large areas of the rock sample, attaining
potentially the number of magnetic moments necessary for paleomagnetic studies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Methodology}

We will achieve the goal of estimating the dipole moment of several individual
grains per image in a semi-automatic fashion by dividing the task into three
parts that can be performed independently:

\begin{enumerate}
\item \textbf{Source detection and separation:} Identify and spatially isolate
  the magnetic field caused by each source. We will do this through a
  combination of classic potential field data processing (\textit{total
  gradient amplitude}) and image processing (histogram stretching and
  equalization) and segmentation methods (watershed segmentation).
\item \textbf{Position estimation}: Estimate the 3D position of a magnetic
  particle based on the magnetic field measurements and the assumption of a
  dipolar source. This can be achieved by applying the Euler Deconvolution
  method to the data segment identified in step 1.
\item \textbf{Magnetic moment inversion:} Estimate the 3-component dipole
  moment vector by inverting the magnetic field data using the position
  obtained from Euler Deconvolution as a constraint and the assumption of a
  dipolar source. This leads to a linear inverse problem that is stable and
  computationally efficient, particularly since the inversion is performed
  separately for each data segment identified in step 1.
\end{enumerate}

In the sections below, we describe the methodology used in each step of our
proposed workflow.

\subsection{Automatic source detection and separation}

Our first task is to automatically identify the signal of individual particles
and spatially segregate the data into windows, each containing the signal of a
single particle. We implemented the following workflow to achieve this:

\begin{enumerate}
\item Calculate the \textit{total gradient amplitude} (TGA), also known as the 3D analytic signal \citep{Roest1992Magnetic}, of the observed vertical   component of the magnetic field. The TGA is entirely positive and tends to be more concentrated on top of the magnetic field sources than the observed magnetic field. The derivative calculation also acts as a high-pass filter, removing long-wavelength noise from the data.
\item Apply a contrast stretching method to re-scale the TGA to a new range   defined by the lower and upper percentiles of the data in order to highlight the weaker signals, either coming from small or from deep-seated particles.
%\item To enhance the quality of an image before the blob detection, we can also apply a common two-step approach that involves using a median filter and an unsharp mask filter \citep{VanderWalt2014}. The median filter reduces high-frequency noise (e.g., salt and pepper noise) in the image at the cost of smoothing slightly the edges, and the unsharp mask filter enhances the contrast and sharpness of edges. By applying a median filter followed by an unsharp mask filter, we can significantly improve the visual quality of an image while preserving the edges and the details.  
\item Use the Laplacian of Gaussian (LoG) method \citep{VanderWalt2014} on the   re-scaled TGA data to estimate the position and size of data windows containing the signal of each particle.
\end{enumerate}

The total gradient amplitude (TGA) was devised as a filter that can be applied
to aeromagnetic data to reduce the signal's dependence on the direction of
magnetization of the source and that concentrates the signal above the sources
\citep{Roest1992Magnetic, Nabighian2005}. The TGA is defined as the norm of the
gradient vector of a scalar field $f(x, y, z)$

\begin{equation}
\|\vec{\mathbf{\nabla}}f(x, y, z)\|  =
\sqrt{(\partial_x f)^2 + (\partial_y f)^2 + (\partial_z f)^2}
\ ,
\end{equation}

\noindent
in which $\partial_x f$ is the partial derivative of $f$ with respect to $x$
and likewise for the $y$ and $z$ directions. The partial derivatives are best
approximated using a second-order accurate central finite-difference scheme

\begin{equation}
\partial_x f(x, y, z) \approx
\dfrac{f(x + \Delta x, y, z) - f(x - \Delta x, y, z)}{2 \Delta x}
\ ,
\end{equation}

\noindent
and likewise for the $y$ and $z$ directions, in which $\Delta x$ is the grid
spacing (assumed to be equal in the $x$ and $y$ directions). For the $z$
component, $\Delta z = \Delta x$ and $f(x, y, z + \Delta z)$ and
$f(x, y, z - \Delta z)$ are calculated by upward and downward continuation,
respectively, performed in the wavenumber domain. The 2D maps of the $x$, $y$,
and $z$ derivatives will also be used in the Euler Deconvolution step described
below, which is known to be highly sensitive to noise in the derivatives
\citep{Saleh2012Applying}. This is why e prefer the finite-differences based
derivatives, which are less prone to amplifying short-wavelength noise than
those calculated in the wavenumber domain through the Fast Fourier Transform
(FFT).

Once we have obtained a TGA image, we apply a contrast stretching method to
re-scale the TGA values in order to highlight the weaker signals present in the
image. This is a necessary step to make sure that the blob detection algorithm
is able to identify particles causing both weak and strong signals. The
following contrast stretching operation is performed per pixel of the TGA image

\begin{equation}
\text{TGA}_{rescaled} =
2\left(\dfrac{\text{TGA} - v_{min}}{v_{max} - v_{min}}\right) - 1
\ ,
\end{equation}

\noindent
in which $v_{min}$ and $v_{max}$ are the upper and lower bounds of TGA values
that will be stretched to the range $[0, 1]$. By experimentation, we found
that values of $v_{min} = 1^\text{st}$ and $v_{max} = 99^\text{th}$
percentiles of the TGA range work well on real magnetic microscopy datasets.

The re-scaled TGA image is then used as input for a Laplacian of Gaussian (LoG)
blob detection algorithm \citep{Kong2013}. This method is able to identify the
location and size of multiple windows in the image containing ``blobs''. In our
case, the blobs are the re-scaled TGA field of each particle and the LoG method
detects the local TGA maxima. The LoG method is particularly well suited for
the detection of bright blobs on a dark background at the expense of a longer
computation time \citep{Han2016}. For the image sizes routinely found in
magnetic microscopy, the computation is fast and only needs to be performed
once per image.

\subsection{Euler Deconvolution}

Once the locations and sizes of the windows containing the isolated signals of
each particle are determined, we can apply the Euler Deconvolution (ED)
estimate of the $x$, $y$, and $z$ positions of the center of the particles
under a dipolar approximation. This technique was first proposed by
\citet{Thompson1982} under the name EULDPH and later extended to three
dimensions and renamed ``Euler Deconvolution'' by \citet{Reid1990}. ED is
traditionally performed on a set of moving windows that scan the entire
dataset, producing a large scatter of position estimates, most of which are
spurious \citep{Silva20033D}. This is only done because it is impractical to
segment an aeromagnetic dataset into individual anomalies. Fortunately,
magnetic microscopy images contain fewer elongated features (e.g., dikes and
suture zones) and regional signals (e.g., Curie depth variations) that are
difficult to separate. This makes it possible for us to produce isolated
signals for each magnetic particle through the source detection step described
above. Since each data window contains only the signal of a single particle, we
are able to apply ED and generate only a single position estimate per particle.

\begin{figure}[tb]
\centering
\includegraphics[width=\linewidth]{figures/coordinate-system-sketch.png}
\caption{
Schematic representation of the coordinates systems and modelling elements.
The left panel shows the $x, y, z$ right-handed coordinate system with $z$
pointing upward and away from the sample. The top surface of the sample defines
the $z=0$ surface. A dipole with dipole moment vector $\mathbf{\vec{m}}$ is
also shown at coordinates $(x_c, y_c, z_c)$. The observation points are located
at regular intervals on an $x-y$ plane at a positive $z$ separation from the
sample surface. The right panel zooms in on the dipole and shows the dipole
moment vector expressed in terms of inclination ($I$, positive downwards),
declination ($D$, angle with the $\hat{y}$ direction), and moment magnitude $m
= \|\mathbf{m}\|$.
}
\label{fig_coordinate_systems}
\end{figure}

Euler Deconvolution is formulated as a least-squares inversion of Euler's
homogeneity equation

\begin{equation}
\label{eq_euler_homogeneity}
(x - x_c)\partial_x f
+ (y - y_c)\partial_y f
+ (z - z_c)\partial_z f
= (b - f)\eta
\ ,
\end{equation}

\noindent
in which $(x_c, y_c, z_c)$ are the coordinates of the magnetic field source (Figure~\ref{fig_coordinate_systems}), $b$ is the base level representing a constant shift in the signal, and $\eta$ is the structural index corresponding to the nature of the source \citep{Reid1990}. This equation holds true for simple geometric sources, like spheres, dipoles, and vertical cylinders. Here, we assume that the magnetic particles are small enough and the sensor is far enough away that the sources can be represented by dipoles, yielding an $\eta=3$.

The inversion is performed by rearranging Equation~\ref{eq_euler_homogeneity} into a pseudo-parametric model with parameters $x_c$, $y_c$, $z_c$, and $b$

\begin{equation}
x_c \partial_x f + y_c \partial_x f + z_c \partial_x f + \eta b
=
x \partial_x f + y \partial_y f + z \partial_z f + \eta f
\ .
\end{equation}

Given a set of $N$ observations of the magnetic field as the harmonic function $f$ and its spatial derivatives, we can form the $N \times 4$ system of equations

\begin{equation}
\begin{bmatrix}
  {\partial_x f}_1 & {\partial_y f}_1 & {\partial_z f}_1 & \eta \\
  {\partial_x f}_2 & {\partial_y f}_2 & {\partial_z f}_2 & \eta \\
  \vdots & \vdots & \vdots & \vdots \\
  {\partial_x f}_N & {\partial_y f}_N & {\partial_z f}_N & \eta
\end{bmatrix}
\begin{bmatrix}
  x_c \\ y_c \\ z_c \\ b
\end{bmatrix}
=
\begin{bmatrix}
  x_1 {\partial_x f}_1 + y_1 {\partial_y f}_1 + z_1 {\partial_z f}_1 + \eta f_1 \\
  x_2 {\partial_x f}_2 + y_2 {\partial_y f}_2 + z_2 {\partial_z f}_2 + \eta f_2 \\
  \vdots \\
  x_N {\partial_x f}_N + y_N {\partial_y f}_N + z_N {\partial_z f}_N + \eta f_N \\
\end{bmatrix}
\ .
\end{equation}

In matrix notation, this linear system can be written as

\begin{equation}
\label{eq_euler_forward}
\mathbf{G} \mathbf{p} = \mathbf{h} \ .
\end{equation}

We can arrive at a solution to Equation~\ref{eq_euler_forward} by assuming that the three spatial derivatives of $f$ have negligible error and minimizing the misfit $\phi(\mathbf{p})$ between a pseudo-observation vector $\mathbf{h}^o$ and the predictions $\mathbf{h}$. The least-squares misfit $\phi(\mathbf{p})$ is defined as

\begin{equation}
\label{ZTSuSBbL16}
\phi(\mathbf{p}) = \|\mathbf{h}^o - \mathbf{h}\|^2 = (\mathbf{h}^o - \mathbf{G}\mathbf{p})^T (\mathbf{h}^o - \mathbf{G}\mathbf{p})\ .
\end{equation}

The minimum of $\phi(\mathbf{p})$ is obtained by solving the $4 \times 4$
system of normal equations

\begin{equation}
\mathbf{G}^T \mathbf{G} \mathbf{p} = \mathbf{G}^T \mathbf{h}^o\ .
\end{equation}

The solution vector $\hat{\mathbf{p}}$ provides an estimate of the position
$(x_c, y_c, z_c)$ and base level $b$ for the source located inside of a data
window. Repeating this process for each window produced by the source detection
algorithm will yield the horizontal locations and depths of each magnetic
particle.

\subsection{Magnetic moment inversion}

Once the source position is known and we can assume that it is approximately
spherical or dipolar, we can apply the method developed by
\citet{Oliveira2015Estimation} to estimate the dipole moment vector
$\mathbf{m}$ of the source. We begin by following
\citet{Oliveira2015Estimation} in formulating the magnetic induction vector
$\mathbf{b}$ of a dipole as

\begin{equation}
\label{eq_vector_dipole_field}
\mathbf{b}
=
\begin{bmatrix}
  b_x \\ b_y \\ b_z
\end{bmatrix}
= \dfrac{\mu_0}{4\pi}
\begin{bmatrix}
    \dfrac{\partial^2}{\partial x \partial x} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial x \partial y} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial x \partial z} \dfrac{1}{r}
  \\
    \dfrac{\partial^2}{\partial y \partial x} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial y \partial y} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial y \partial z} \dfrac{1}{r}
  \\
  \dfrac{\partial^2}{\partial z \partial x} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial z \partial y} \dfrac{1}{r}
  & \dfrac{\partial^2}{\partial z \partial z} \dfrac{1}{r}
\end{bmatrix}
\begin{bmatrix}
  m_x \\ m_y \\ m_z
\end{bmatrix}
= \dfrac{\mu_0}{4\pi} \mathbf{M}\,\mathbf{m}
\ ,
\end{equation}

\noindent
in which $r = \sqrt{(x - x_c)^2 + (y - y_c)^2 + (z - z_c)^2}$ is the Cartesian distance between the observation point $(x, y, z)$ and the source $(x_c, y_c, z_c)$ and $\mu_0$ is the vacuum magnetic permeability. Most magnetic microscopes provide measurements of only the vertical component $b_z$, which can be isolated from Equation~\ref{eq_vector_dipole_field} as shown in Equation~\ref{eq_dipole_bz}, which is a similar approach to the uniform model proposed by \cite{Weiss2007}.

\begin{equation}
\label{eq_dipole_bz}
b_z
= \dfrac{\mu_0}{4\pi}
\begin{bmatrix}
\dfrac{\partial^2}{\partial z \partial x} \dfrac{1}{r}
& \dfrac{\partial^2}{\partial z \partial y} \dfrac{1}{r}
& \dfrac{\partial^2}{\partial z \partial z} \dfrac{1}{r}
\end{bmatrix}
\begin{bmatrix}
m_x \\ m_y \\ m_z
\end{bmatrix}
= \dfrac{\mu_0}{4\pi} \mathbf{M_z}\,\mathbf{m}
\ .
\end{equation}

The three second-order derivatives in Equation~\ref{eq_dipole_bz} are

\begin{equation}
\begin{aligned}
\dfrac{\partial^2}{\partial z \partial x} \dfrac{1}{r} &=
\dfrac{3(z - z_c)(x - x_c)}{r^5}\ ,
\\
\dfrac{\partial^2}{\partial z \partial y} \dfrac{1}{r} &=
\dfrac{3(z - z_c)(y - y_c)}{r^5}\ ,
\\
\dfrac{\partial^2}{\partial z \partial z} \dfrac{1}{r} &=
\dfrac{3(z - z_c)^2}{r^5} - \dfrac{1}{r^3}\ .
\end{aligned}
\end{equation}

Given a set of $N$ observations of $b_z$ made inside a window containing a single source, we can form the $N \times 3$ linear equation system

\begin{equation}
\label{CgjOtKLQKT}
\begin{bmatrix}
\dfrac{\mu_0}{4\pi}\dfrac{3(z_1 - z_c)(x_1 - x_c)}{r_1^5}
& \dfrac{\mu_0}{4\pi}\dfrac{3(z_1 - z_c)(y_1 - y_c)}{r_1^5}
& \dfrac{\mu_0}{4\pi}\left(\dfrac{3(z_1 - z_c)^2}{r_1^5} - \dfrac{1}{r_1^3}\right)
\\
\dfrac{\mu_0}{4\pi}\dfrac{3(z_2 - z_c)(x_2 - x_c)}{r_2^5}
& \dfrac{\mu_0}{4\pi}\dfrac{3(z_2 - z_c)(y_2 - y_c)}{r_2^5}
& \dfrac{\mu_0}{4\pi}\left(\dfrac{3(z_2 - z_c)^2}{r_2^5} - \dfrac{1}{r_2^3}\right)
\\
\vdots & \vdots & \vdots
\\
\dfrac{\mu_0}{4\pi}\dfrac{3(z_N - z_c)(x_N - x_c)}{r_N^5}
& \dfrac{\mu_0}{4\pi}\dfrac{3(z_N - z_c)(y_N - y_c)}{r_N^5}
& \dfrac{\mu_0}{4\pi}\left(\dfrac{3(z_N - z_c)^2}{r_N^5} - \dfrac{1}{r_N^3}\right)
\end{bmatrix}
\begin{bmatrix}
m_x \\ m_y \\ m_z
\end{bmatrix}
=
\begin{bmatrix}
{b_z}_1 \\ {b_z}_2 \\ \vdots \\ {b_z}_L
\end{bmatrix}
\ ,
\end{equation}

\noindent
which can also be expressed in matrix form as

\begin{equation}
\label{qdhqM4s9Ln}
\mathbf{A} \mathbf{m} = \mathbf{d} \ .
\end{equation}

As with Euler Deconvolution, we can find a dipole moment vector that best fits
a set of $N$ observations of the vertical component of the magnetic field
$\mathbf{d}^o$ in a least-squares sense by minimizing the misfit function

\begin{equation}
\label{uV9pRVYO4l}
\Gamma (\mathbf{m}) = \| \mathbf{d}^o - \mathbf{d} \|^2 = (\mathbf{d}^o - \mathbf{A}\mathbf{m})^T  (\mathbf{d}^o - \mathbf{A}\mathbf{m})\ .
\end{equation}

\noindent
The dipole moment vector that minimizes $\Gamma (\mathbf{m})$ can be found by
solving the $3 \times 3$ normal equation system

\begin{equation}
\mathbf{A}^T \mathbf{A} \mathbf{m} = \mathbf{A}^T\mathbf{d}^o\ .
\end{equation}

\noindent
The estimated dipole moment vector $\hat{\mathbf{m}}$ can be converted into declination $D$, inclination $I$, and magnitude $m$, which are more intuitive quantities for interpreting paleomagnetic results, like so

\begin{equation}
  \label{eq_vector_to_incdec}
\begin{aligned}
I &= \tan^{-1}\left(\dfrac{m_z}{\sqrt{m_x^2 + m_y^2}}\right) \ , \\
D &= \tan^{-1}\left(\dfrac{m_x}{m_y}\right) \ , \\
m &= \sqrt{m_x^2 + m_y^2 + m_z^2} \ .
\end{aligned}
\end{equation}

It is important to note that these are not paleomagnetic declination and inclination angles, but are instead related to the sample coordinate system. To obtain the paleomagnetic direction, the dipole moment vector must be rotated to the sample field orientation prior to the application of Equation~\ref{eq_vector_to_incdec}.

In a micromagnetic survey, or any geophysical survey, measurements are affected by noise caused by experimental errors, equipment inaccuracies, and sensor limitations. This noise added to the observed data vector affects the solution of the estimated parameter vector $\mathbf{m}$, regardless of the method used. Assuming that the noise in the observed data measure is independent and of variance ${\sigma_0}^2$, the latter given by (i) the squared standard deviation used in synthetic cases (e.g., pseudo-random normal distributions) or (ii) the sensor limitations for real samples. Nevertheless, if the sample's variance is not known, one can typically estimate it by using the \emph{$\chi^2$} statistic, \citep[Eq.~\ref{chi_square},][]{Aster2019}, which depends on the residual sum of squares and the degrees of freedom (\emph{v}), the latter being the difference between the number of observed data points (N) and the parameters (m).

\begin{equation}
  \label{chi_square}
\begin{aligned}
\large \sigma_0^2 = {\chi}^2 = \frac{\sum{\left( b_z - {b_{z}}_{dipole} \right)}^2}{N-m}
\end{aligned}
\end{equation}

\begin{equation}
\label{lPOQaqlV3N}
\mathbf{C} =
\begin{bmatrix}
\sigma_x^2 & \sigma_{xy} & \sigma_{xz} \\
\sigma_{yx} & \sigma_y^2 & \sigma_{yz} \\
\sigma_{zx} & \sigma_{zy} & \sigma_z^2
\end{bmatrix}
=
\sigma_0^2 (\mathbf{A}^T\mathbf{A})^{-1}
\end{equation}

From the main diagonal of $\mathbf{C}$ we can obtain the variances of the
declination $\sigma_D^2$, inclination $\sigma_I^2$, and magnitude $\sigma_m^2$
by the propagation of uncertainties

\begin{equation}
\label{uncertanties}
\begin{aligned}
\sigma_D^2 &= \dfrac{m_y^2\sigma_x^2 + m_x^2\sigma_y^2}{\left(m_x^2 + m_y^2\right)^2} \ , \\
\sigma_I^2 &= \dfrac{m_x^2 m_z^2 \sigma_x^2 + m_y^2 m_z^2 \sigma_y^2 + \left(m_x^2 + m_y^2\right)^2\sigma_z^2}{\left(m_x^2 + m_y^2\right) m^4} \ , \\
\sigma_m^2 &= \dfrac{m_x^2\sigma_x^2 + m_y^2\sigma_y^2 + m_z^2\sigma_z^2}{m^2} \ .
\end{aligned}
\end{equation}

\noindent
These variances reflect the sensitivity of the estimated dipole moment to random noise in the magnetic field observations. However, they do not capture other, often larger, sources of uncertainty like systematic errors in the observations, data positioning errors, and the validity of the dipolar approximation. Therefore, we recommend that these estimated variances be treated with caution and should not be interpreted as ``the degree of certainty that the estimated values are the true values''.

It is relevant to highlight that parameters retrieved during inversion procedures do not always explain the observed data, so it is necessary to evaluate the fitting of the predicted data obtained with these parameters. \cite{Fu2020} apply the calculation of the "non-dipolarity parameter (\emph{D})" by calculating the coefficient of determination ($R^2$, Eq.~\ref{R2}). The latter has a maximum value of 1, if \emph{D} is close enough to this value, a simple dipole model is able to explain the measured data. On the other hand, low values of $R^2$ indicate that the predicted model does not explain the observed data. \cite{CortesOrtuno2021} use another approach for the evaluation of the predicted model fit, called signal-to-noise ratio (SNR), here we modified it to be given in decibels (Eq.~\ref{SNR}). SNR is the trade-off between the signal and its associated noise acquired during the measurement by any sensor. The higher the SNR values the better it will be the fit of the predicted model. For a signal to be “visible”, \cite{Strum2014} suggest an SNR value of at least 3, which means the signal is 3 times bigger than the noise. Thus, both $R^2$ and the concept of SNR are appropriate to be used as criteria filtering for reliability for dipolar inversions.

\begin{equation}
\label{R2}
\begin{aligned}
\large R^2 = 1 - \frac{ \sum{\left( b_z - {b_{z}}_{dipole} \right)}^2 }{\sum{\left( b_z - \bar{b_z} \right)}^2}
\end{aligned}
\end{equation}

\begin{equation}
\label{SNR}
\begin{aligned}
\large SNR = 10 \log_{10}{\left( \frac{{{\sigma}^2}_{dipole}}{{{\sigma}^2}_{residual}} \right)}
\end{aligned}
\end{equation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to synthetic data}

We first applied our inversion workflow to synthetic data to evaluate its strengths and limitations. The first set of synthetic data is simple and intended to test the method's performance under ideal circumstances. The second set is more complex, with a greater number of magnetic particles and long-wavelength noise, and is intended to demonstrate the capabilities and limitations of the method under real-world scenarios.

The first synthetic model was performed by generating the vertical magnetic field derived from a simple set of four spheres with similar magnetization intensity and radius, but completely different magnetization directions, contaminated by high-frequency noise. This simplified test model is basically done to investigate the efficiency of the inversions used for Euler positioning and magnetization parameters, thus being a validation of the methodology.

The second model contains one hundred spheres with different magnetization intensities magnetized in two stable directions and similar standard deviation. The magnetic field generated from these sources is corrupted by both low and high-frequency noise. The complexity of this synthetic data seeks to more faithfully emulate real magnetic microscopy data.

\subsection{Simple simulation: Validating the methodology}

\begin{figure}[tb!]
  \centering
  \includegraphics[width=1\linewidth]{figures/simple-synthetic-data.png}
  \caption{
    Synthetic data and the various processing steps performed prior to the
    dipole moment inversion. 
    a) The synthetic noise-corrupted $b_z$ observations at 
    $z = \qty{5}{\micro\meter}$ due to four dipolar sources with different
    depths and dipole moment vectors 
    (see Table~\ref{tab_synthetic_simple_results}).
    b) The upward-continued data to $z = \qty{10}{\micro\meter}$ showing
    attenuated short-wavelength noise.
    c) The total gradient amplitude (TGA) calculated from the 
    upward-continued data, which is able to concentrate the signal on top
    of each dipolar source. 
    d) The contrast-stretched TGA, highlighting the signal of all four
    sources, including the weak central source.
    e) The detected source bounding boxes (black squares) that correctly
    identify the signal of all four sources.
    f) The estimated source locations (colored circles) from Euler 
    Deconvolution of the upward-continued data inside each bounding box. 
    The color represents the difference between the true and estimated 
    $z$ coordinates.
  }
  \label{fig_synthetic_simple_data}
\end{figure}

The Figure~\ref{fig_synthetic_simple_data}a shows the vertical component of the magnetic field $b_z$ over a synthetic rock section containing four uniformly magnetized spherical sources. The map covers a surface of $\qty{1000}{\um} \times \qty{1000}{\um}$ with data points in a regular grid of approximate \qty{1}{\um} ($N = 10^6$ observations) obtained at a sensor-sample distance of \qty{5}{\um}. The magnetic field data is contaminated with a pseudo-random noise of normal Gaussian distribution with a zero mean and \qty{20}{\nano\tesla} standard deviation.

We first applied an upward continuation filter (Fig.~\ref{fig_synthetic_simple_data}b) to smooth out high-frequency noise \citep{Blakely1996}, since it strongly affects the first derivatives of the potential field required by the Euler equation. Then, we calculated the magnetic total gradient anomaly map (Fig.~\ref{fig_synthetic_simple_data}c), which is subjected to a stretch contrast filter to highlight as much as possible the weaker intensity sources (Fig.~\ref{fig_synthetic_simple_data}d), 
in order to facilitate the next step. Subsequently, we applied the blob detection algorithm to the contrasted total gradient map to obtain the position of the data window for each source (Fig.~\ref{fig_synthetic_simple_data}e). The ED was performed for each selected source by assuming the structural index of a sphere/punctual source, therefore obtaining the Cartesian coordinates (Fig.~\ref{fig_synthetic_simple_data}f) of each source causing the magnetic potential anomaly in the observed data set (Table~\ref{tab_synthetic_simple_results}).

\begin{figure}[tb!]
\centering
\includegraphics[width=0.5\linewidth]{figures/simple-synthetic-dipole-moment.png}
\caption{
Comparison of true and estimated dipole magnetic moments and directions.
}
\label{fig_synthetic_simple_results}
\end{figure}

The central positions of each source obtained with ED were used as an input parameter for inverting the vertical magnetic field. The inversion produced a least squares approximation solution for the vector of Cartesian magnetic moments $m_x, m_y, m_z$ and the corresponding magnetization directions (D and I) and intensity of magnetic moment $m$ (Fig.~\ref{fig_synthetic_simple_results}). We also obtained the uncertainty propagation of this inversion, using equations~\ref{chi_square}-\ref{uncertanties}, as shown in Table~\ref{tab_synthetic_simple_results}.

\begin{table}[t]
  \begin{center}
    \small
    \input{simple-synthetic-results.tex}
  \end{center}
  \caption{Each source positioning and magnetization parameters modeled and recovered by the least square estimator, respectively.}
  \label{tab_synthetic_simple_results}
\end{table}

\subsection{Complex simulation: Testing applicability}

This test represents a more complex scenario by simulating 100 sources randomly distributed in the imaged area of a synthetic thin section of 2000 x 2000 $\mu m$. The synthetic $b_z$ data were generated on a regular grid with 2 $\mu m$ spacing and 5 $\mu m$ sample-sensor distance. For greater fidelity to real samples, the magnetic sources are modeled as dipoles with different depths and magnetic moment intensity. The depths vary between 1 and 20 $\mu m$, while the magnetic moment intensity ranges from $10^{-12}$ to $10^{-14}$ $Am^2$.

To add more complexity, we simulate two stable magnetization  directions, both sampled from a pseudo-random Gaussian distribution, this randomness simulates the NRM found in real ferromagnetic particles that vary individually but average out to the inducing field direction. The first group of sources ($n = 70$) has a magnetization mean direction of D = 0°, I = 0°, and a standard deviation of 10°. While the second one ($n = 30$) has a magnetization mean direction of D = 180°, I = 0°, also with standard deviation of 10°. Finally, we contaminated the data with high-frequency normally-distributed pseudo-random noise with zero mean and 50 nT standard deviation, as well as with low-frequency noise in the form of additional deep sources beyond the modeling domain. The noise-corrupted synthetic $b_z$ image is shown in Figure~\ref{complex-synthetic-data}a.

\begin{figure}[t]
  \centering
  \includegraphics[width=1\linewidth]{figures/complex-synthetic-data.png}
  \caption{
    a) The synthetic high and low-frequency noise-corrupted $b_z$ observations at 
    $z = \qty{5}{\micro\meter}$ due to two clusters of stable directions simulated. 
    b) Residual anomaly map after the $5^{th}$ degree polynomial signal subtraction and upward-continued data to $z = \qty{10}{\micro\meter}$ to attenuated long and short-wavelength noise, respectively.
    c) The total gradient amplitude (TGA) calculated from the 
    upward-continued data, which is able to concentrate the signal on top
    of each dipolar source. 
    d) The contrast-stretched TGA, highlighting the signal of all sources, especially the weaker ones.
    e) The detected source bounding boxes (black squares) that correctly
    encapsulate the signal of the sources.
    f) The estimated source locations (colored circles) from Euler 
    Deconvolution of the upward-continued data inside each bounding box. 
    The color represents the difference between the true and estimated 
    $z$ coordinates.
  }
  \label{complex-synthetic-data}
\end{figure}

Since now the data have both high and low-frequency noise, we performed a $5^{th}$ degree polynomial fitting to remove the low-frequency wavelength from the signal. Subsequently, we applied an upward continuation to get rid of the high-frequency ones (Fig.~\ref{complex-synthetic-data}b). Followed by the TGA map calculation (Fig.~\ref{complex-synthetic-data}c), stretched-contrast filter (Fig.~\ref{complex-synthetic-data}d), and blob detection algorithm (Fig.~\ref{complex-synthetic-data}e). The positioning result of each source by window detection method is shown in Figure~\ref{complex-synthetic-data}e, where 98 particles out of the original 100 were detected.

The spatiality of the inversion results is one of the key advantages of magnetic microscopy over the classic techniques of paleomagnetism. Therefore, we present the inversion results spatially so that we can evaluate any patterns in their distribution. The depth and dipolar moment of each particle are represented by horizontal and vertical error bars (Fig.~\ref{complex-synthetic-comparison}), respectively. These two variables are useful proxies for reliably determining the range for the width and amplitude range of the vertical component magnetic signal the technique is able to identify, which is crucial to know when evaluating these types of inversion results.

Figure~\ref{complex-synthetic-comparison} shows the difference between the estimated dipole moments and the true values for each of the 98 particles that were identified and the model fitting tests. Firstly, we evaluate the magnetic direction misfits by calculating the module of the angular difference between the true and the estimated magnetization vectors (Fig.~\ref{complex-synthetic-comparison}a). Afterward, we calculate the dipole moment misfits in percentage (~$|~100 \left( m_{true} - m_{estimate}\right)~/~\left(m_{true}\right)~|$~), as shown in Figure~\ref{complex-synthetic-comparison}b). Then, the fitting tests were performed using the equations~\ref{R2}-~\ref{SNR} in an attempt to select which criterion best separates data with good fitting and low errors of direction and dipole moment. Figure~\ref{complex-synthetic-comparison}c shows the R² coefficient, or the \emph{D} parameter of the non-dipolarity test \citep{Fu2020}, representing how well the dipolar model is able to explain the observed data in the given window. On the other hand, the SNR is capable to express the power of the predicted model over its residual, therefore can select which sources are more "visible" by the sensor (Fig.~\ref{complex-synthetic-comparison}d). Consequently yielding results directly proportional to the R² coefficient.

\begin{figure}[htbp!]
\centering
\includegraphics[width=0.75\linewidth]{figures/complex-synthetic-comparison.png}
\caption{
The validation of the result obtained with the inversion was calculated for each individual particle based on the error between the real parameters modeled and their respective recovered values, being (a) the direction and (b) intensity of the magnetic moment, in addition to the $R^2$ score (c) obtained by comparing the forward model and the actual data. The depth and radius of the magnetic sources are also important factors that influence the final result, therefore, these data are given in the form of cross plots, with the vertical bar represented by the depth (1 - 20 $\mu m$) and the horizontal bar by the dipole moment amplitude ($10^{-12}$ to $10^{-14}$ $Am^2$).
}
\label{complex-synthetic-comparison}
\end{figure}

Figure~\ref{complex-synthetic-stereograms} shows stereograms with the directions generated by the modeled (Fig.~\ref{complex-synthetic-stereograms}a) and the estimated vectors (Fig.~\ref{complex-synthetic-stereograms}b) for each source. The distribution of the estimated directions is coincident with the true directions aside from a few sources, the same ones with the higher values of direction misfit (Fig.~\ref{complex-synthetic-comparison}a) which is probably associated with the mutual interference of sources close to each other or within the same window. When filtered to include only data with R\textsuperscript{2} \textgreater 0.95 and SNR \textgreater 10 the obtained direction distribution is closer to the true distribution (Fig.~\ref{complex-synthetic-stereograms}c).

\begin{figure}[htbp!]
\centering
\includegraphics[width=1\linewidth]{figures/complex-synthetic-stereograms.png}
\caption{
a) Simulation of a thin section of rock with two sets of stable particles uniformly magnetized (dipolar sources) in the average direction of the induced field (0{\textbackslash}textdegree/0{\textbackslash}textdegree and 180{\textbackslash}textdegree/0{\textbackslash}textdegree. b) The estimated vector directions without filtering. c) Filtering criteria to select the sources' magnetic directions with the best fitting determined by the coefficient of determination ($R^ 2 \geq 0.85$) and signal-to-noise ratio (SNR $ \geq 5$).
}
\label{complex-synthetic-stereograms}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Application to real data}

To test whether the proposed method would be able to determine the
magnetization directions and the magnetic moment of particles in natural
samples, we selected a previously studied carbonate stalagmite sample from the
Wintimdouine cave in the Agadir region (Morocco) \citep{Ait2019Hydro}. This
speleothem contains both magnetite and hematite as the main carriers of
magnetic remanence, as attested by thermomagnetic curves with temperature
decays at {\textasciitilde}580 °C and {\textasciitilde}680 °C, and bimodal
curves of isothermal remanent magnetization (IRM) acquisition \citep{carmo2019speleothem}.

In order to provide two distinct directions associated with the different
magnetic mineral types in this sample, we applied two IRM pulse fields of 2.7 T
and 0.3 T, respectively toward the +Y and -Y directions. In this way, the high
coercivity grains (hematite) would point towards +Y and the low coercivity ones
(magnetite) would align in the -Y direction.

After remanence acquisition, we performed a magnetic map with the Quantum
Diamond Microscope (QDM) at Harvard University over a sample section of
approximately 1410 $\mu$m x 2256 $\mu$m (Figure~\ref{real-data}a) with a grid
spacing of 2.35 $\mu$m and a sensor-sample distance of 5 $\mu$m, totaling
576x10\textsuperscript{3} data points. The QDM is housed in a shielded room, in
order to avoid the influence of the Earth's magnetic field \citep{Fu2020,
Glenn2017}. After applying the magnetic anomaly detection algorithm, it was
possible to determine the windows for 103 potential sources
(Figure~\ref{real-data}e).

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{figures/real-data.png}
\caption{
    a) The real sample vertical component magnetic data $b_z$ observations at 
    $z = \qty{5}{\micro\meter}$. 
    b) Residual anomaly map after the $5^{th}$ degree polynomial signal subtraction and upward-continued data to $z = \qty{10}{\micro\meter}$ to attenuated long and short-wavelength noise, respectively.
    c) The total gradient amplitude (TGA) calculated from the 
    upward-continued data. 
    d) The contrast-stretched TGA.
    e) The detected source bounding boxes (black squares) that correctly
    encapsulate the main signal of the sources.
    f) The estimated source locations (colored circles) from Euler 
    Deconvolution of the upward-continued data inside each bounding box. 
    The color represents the estimated $z$ coordinates.
  }
\label{real-data}
\end{figure}

After applying the ED algorithm, we performed the inversion of the magnetic moment for each of the 103 windows selected previously. In order to reduce considerably the computation time, the inversions were done within each data window, instead of solving all sources parameters at the same time. We obtained the magnetic moment and the direction for all 103 magnetic grains (Figure~\ref{real-data-stereograms}a). Then, we calculated the residuals of the inversions in each window and subsequently the coefficient of determination R\textsuperscript{2} and signal-to-noise ratio (SNR), which were parameters used as filters for the best directions obtained. We considered a good fit to be achieved when $R^2$ was greater or equal to 0.85 and SNR was greater than 5. An $R^2$ value of 0.85 indicates that 85\% of the measured magnetic data is explained by the predicted dipole model, which is equivalent as the non-dipolarity test approach \citep{Fu2020}. On the other hand, an SNR of 5 means that the dipolar signal is 5 times greater than the residual noise. By using these criteria, we ensured that the fit was sufficiently accurate and the dipole model provided a reliable approximation of the original magnetic data. This filtering technique removed the poorly fit predicted models as well as the ones too corrupted with noise, showing more clearly the expected directional clusters of hematite and magnetite crystals (Figure~\ref{real-data-stereograms}b).

\begin{figure}[tb]
\centering
\includegraphics[width=1\linewidth]{figures/real-data-stereograms.png}
\caption{
a) The estimated directions without filtering. b) Estimated directions filtered by the coefficient of determination ($\geq 0.85$) and SNR ($\geq 5$), which shows two clusters of direction located on each pole of the stereogram.
}
\label{real-data-stereograms}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Discussion}

\subsection{Prior information and uniqueness of solutions}
% simple synthetic
Working with potential field data might be very tricky once there is too much ambiguity involved during data modeling, thus to obtain unique and reliable results from data inversion it is necessary to provide as much prior information as possible. One way to circumvent the ambiguity is incorporate prior information about the subsurface structure, such as the positioning of a known source of magnetization. Magnetic field measurements are more sensitive to changes in magnetization near the source that is causing the anomaly. When the position of the source is known, it constrains the model to be as consistent as possible to the observed data, hence increasing the likelihood of obtaining unique solutions. \citet{Oliveira2015Estimation} proved that the magnetization directions (D and I) recovered by the least squares estimator are sensitive to variations in the horizontal coordinates of the center of the magnetic sources, but are practically insensitive to variations in depth. Thus, they consider the ED method as an adequate technique to estimate the central positions that will be used as prior information for inversion. This occurs mainly because, when well performed, the recovery of the source's horizontal coordinates is considerably accurate, while the vertical coordinate can undergo greater variation even though it still provides satisfactory results \citep{Silva20033D, Melo2013}. This remark is also observed with the our simple synthetic samples where the horizontal positions estimated slightly deviate from the true values, which implies very little misfits in the recovered magnetic directions. Although the estimated magnetic moment for the said sample is satisfactory, this magnetic parameter is more affected by the variations in the source's depth, which is cause by the ambiguity mentioned before. In summary, in order to estimate all magnetic parameters as precisely as possible the ED must be executed within a data window containing the lowest amount possible of noise, since that high-frequency noise sensibility is one of the main limitations of the technique.

%complex synthetic
The application in complex synthetic data allows observing the strengths and limitations of the proposed methodology. Some of the strengths that can be mentioned are: (i) the applied technique not only detects most of the modeled sources but also (ii) most of the recovered magnetic parameters have considerably low errors, especially in the directions, usually less than 5° (Fig.~\ref{complex-synthetic-comparison}a). (iii) The magnetic moments obtained from well-individualized sources tend to not deviate much from the real values (Fig.~\ref{complex-synthetic-comparison}b) when R\textsuperscript{2} and SNR scores are high (e.g., $\geq 0.95$ and $\geq 10$, respectively) (Fig.~\ref{complex-synthetic-comparison}c-d). (iv) Shallow particles grouped in clusters are usually well individualized during window selection, as well as (v) some isolated particles that produce a weak magnetic signal. The main limitations observed were: (i) the blob detection fails when there are sources too close, grouping them into the same window, thus causing an erroneous result both for Euler deconvolution and for the magnetic parameters. (ii) The very same occurs when there is a source under another, in this case, the magnetic signal is the sum of both. (iii) In clusters of larger and/or deeper particles, although the method individualizes them well, the magnetic signal of the neighboring particles can considerably influence the result of the inversion. As expected, there is a direct relationship between the dipole intensity and depth with the observed errors. It is clear from the error bars in Figure~\ref{complex-synthetic-comparison} that deep particles and/or particles with small magnetic intensity generate worse results essentially because they will produce smaller anomalous fields in the magnetic maps. Note nonetheless, that even sources with small dipole moments when close to the surface are adequately modeled by our method, because of the trade-off between signal and noise for grains near the sensor.

\subsection{Reliability of dipole moment approximation}
% real samples
Our approach relies in the premise of assuming the magnetic anomaly within 
the data window is a response of a dipolar source. The latter is true when working 
with particle signals in the SD magnetic domain state, since they are uniform magnetized 
particles with strong dipolar anomaly. However, \cite{Nagy2017} reported
that particles within the PSD domain can record the paleomagnetic field for longer (than SD ones) periods
of time, being the stabler and also with strongly non-dipolar characteristics. 
Therefore application of the proposed algorithm to natural samples should fail for those PSD
particles. \cite{CortesOrtuno2022} give important insights about the matter, they 
showed that PSD state particles present more accurate inversion results when considering the non-dipole components for
small sample-sensor distances (\textless 1 $\mu$m), but for larger sensor
distances the dipole as approximations are remarkably accurate, as the
higher-order moments decay rapidly with distance and therefore have less
influence on the particle's magnetic signal. Thus, our approach can be
considered reasonable for working with both particle SD and PSD states signals. 
The latter happens mainly due to the sensor height being
usually greater than 5 $\mu m$, considering a particle on the immediate
surface of the sample the higher-order moments are already quite attenuated, thus 
circumventing the prominent problems described for non-dipolar components.

Despite the excellent signal/noise ratio that the SMM provided with the
proximity of the sensor to the sample, it is worth mentioning that the
measurement noise can still overshadow the responses of very weak/small, or
deep particles, generating unreliable inversion results. Therefore, it is
necessary to keep a check to determine if the inversion reached a satisfactory
result, such as the coefficient of determination and the
signal-to-noise ratio suggested by \citep{CortesOrtuno2021}.

While the windows approach violates the fundamental theory of the inversion problem that 
requires the sampled area to be finite and encapsulated by the inversion domain to ensure 
the uniqueness of results \citep{Baratchart2013,Lima2013}, our technique is similar to the 
one reported by \cite{Weiss2007}. The last-mentioned involves thresholding the long-distance interaction 
of each dipole in the SMM data, which excludes the effect of other dipoles by setting their 
contribution to zero, resulting in a sparse matrix that permits faster calculations. In contrast 
to this approach, we employ the TGA map to select the windows and isolate the area containing 
the main signal of the desired dipole, while the area out of the windows boundaries is less sensible to 
variation in magnetic parameters of this particular source, hence we exclude them from the inversion 
domain. This technique allows us to obtain the 3D positioning and an approximation of the dipole moment 
components of more than a hundred sources within less than two seconds.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Conclusion}

We developed an efficient semi-automated method to determine the direction of
magnetization of dipolar sources on a microscale, as well as the estimate of
their magnetic moment. Being ideal for a reinterpretation for the application
of methods of paleomagnetic studies using thin sections of rock samples. This
would be an attempt to improve the quality of results obtained by isolating the
responses of more reliable recorders of the Earth's geomagnetic field.

We also present a new, faster, and cleaner way to solve the Euler equation in
determining the positioning of magnetic anomaly sources using a pre-selection
of magnetic anomaly source windows based on the Laplacian of Gaussian
applied to total gradient anomaly maps. In this way, reducing the numerous
solutions to just one data window per source. After estimating the structural
index ($n = 3$) by approximating the sources generating the magnetic anomaly to
spheres/points, the Euler deconvolution is performed, and the central position
of each source is determined. Due to the similarity with aeromagnetic data, this
approach can also be extrapolated for macro-scale studies.

To recover magnetic direction and moment we only need to assume that the
sources have their central positions known (so we apply Euler deconvolution)
and that their magnetization are uniform. This last premise aligns with the
theory of magnetically stable particles SD (and the PSD ones at a reasonable
sensor height), which is the basis of classical paleomagnetism. Also, there is
no need for any kind of prior knowledge other than the observed magnetic
anomaly, and the structural index of the sources. Therefore, this method can be
quickly replicated in a dataset of thin sections of rocks to obtain the
distributions of magnetic directions of each source identified in the sample.

The test using a simple synthetic sample shows the great capability of the method by retrieving not only the precise center positions, but also retrieving the magnetization directions and intensity even under the considerable effect of high-frequency noise. While the complex synthetic sample data allows observing the applicability of the method developed in real samples that are more complex with varied magnetization directions and intensity, in addition to also taking into account the high and low-frequency noise and sources with variable sizes and depths. The real sample data positively answered the question of the algorithm's ability to deal with rocks thin sections. But also, showed the acceptable capacity of retrieving different magnetization directions recorded by magnetic minerals with different coercivities and magnetic signal disparities even greater than predicted in the complex synthetic test. We also assessed the quality of the fit between the predicted dipole model and the original magnetic data using two criteria: the coefficient of determination ($R^2$) and the signal-to-noise ratio (SNR).

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Data and code availability}

The Python source code used to produce all results and figures presented here
is available at \url{https://github.com/\GitHubRepository} and
\url{https://doi.org/\ArchiveRepository} under the MIT open-source license.

The image re-scaling and blob detection though the Laplacian of Gaussian method
were performed with the scikit-image library \citep{VanderWalt2014}. We also used Matplotlib \citep{Hunter2007} for plotting the figures, numpy \citep{Harris2020} and scipy \citep{2020SciPy-NMeth} for all performed calculations. Verde \citep{verde2018} was used to generate data grids and to perform the regional residual separation, while the upward continuation filter was calculated using Harmonica \citep{harmonica2020}. The Choclo library \citep{choclo2022} provided kernel functions utilized in the forward and inverse problems. Loops were parallelized with the aid of Numba just-in-time compilation \citep{lam2015numba}. Lastly, the xarray library \citep{hoyer2017xarray} offered a fast and powerful tool to work with multi-dimensional datasets allowing an easy way to alteration, visualization and extraction of data with advanced indexing techniques.

The QDM image data is available at CITE A DATAVERSE/FIGSHARE ARCHIVE under the
CC-XX license.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
This research project was supported by the Conselho Nacional de Desenvolvimento Científico e Tecnológico (CNPq) - grant 162704/2021-6 and by the Fundação de Amparo à Pesquisa do Estado de São Paulo (FAPESP) - grant 2021/08379-5, the opinions, hypotheses and conclusions or recommendations expressed in this material are the responsibility of the authors and do not necessarily reflect the views of FAPESP. We also acknowledge the support of the Royal Society \textcolor{red}{[insert grant number]} for this research project. We are indebted to the developers and maintainers of the open-source software without which this work would not have been possible.